Evidence of widespread maintenance of functional composition in vertebrate communities

### 2021-09-16 paper/manuscript.Rmd
Minor reference and grammar changes

#### *Data*

@@

```{r echo = FALSE, results = "asis"}
tibble(Taxa = c("Mammals", "Birds", "Amphibians"),
       Timeseries_Count = c(48, 2380, 11),
       Species_Count = c(184, 700, 184),
       Trait_Source = c("Elton Traits", "Elton Traits", "Amphibio"),
       Traits = c("body mass, diet, active diel period", "body mass, diet, nocturnality, \nforest foraging strata, pelagic specialist", "habitat, diet, active diel period, \nactivity seasonality, body mass, body length, \nmin maturation size, max maturation size, \nmin offspring size, max offspring size, \nreproductive output, breeding strategy")
) %>%
  mutate_all(linebreak, align = "l") %>% 
  knitr::kable(format = "latex", col.names = c("Taxa", "Number of \nTimeseries", "Number of \nSpecies", "Trait Source", "Traits"), escape = FALSE, 
               caption = "Summary of the data in the final trait database.") %>%
  kable_classic() #%>% 
  #kable_styling(latex_options= c("scale_down", "hold_position")) #%>%
   #kableExtra::column_spec(2, width = "5cm") #%>%
  # kableExtra::column_spec(4, width = "10cm")
```

@@

#### *Diversity Metrics*

@@

#### *Null Models*

@@

#### *Analysis*

@@

*Evidence of Ecological Processes*

@@

*Potential Methodological Limitations*

Here we approach the question of functional change using the best available data and biodiversity synthesis approaches.
However, a number of gaps in best practices may be obscuring a true underlying trend.

First, the Biotime database is not a representative sample of the world's biodiversity or areas of greatest threat [@gonzalez2016; @vellend2017], and the subset of data in this study exhibits a strong Northern Hemisphere bias. We may simply not have data from those areas experiencing the greatest perturbation [@hughes2021]. We do however emphasize that even disturbed communities maintain functional structure [@edwards2013; @matuoka2020], and the inclusion of more heavily impacted sites would not necessarily change the result.

@@

Third, while we use here the most common metrics for describing functional diversity they do not measure some potentially important aspects of the functional space.
Most notably, the metrics we calculated do not capture shifts in the location of the functional space.
For example, two communities could have very similar metric values but no overlap in their trait spaces.
This is especially relevant in the context of biodiversity change as a species loss could be replaced by a species with very different functional attributes, but the replacement would go undetected if the new species expanded the trait space by the same degree and had similar abundance.
This scenario may be common in communities tracking changing environmental conditions.
Approaches for assessing shifts and overlaps in functional space are still relatively new [@barros2016; @blonder2018; @mammola2019] a could shed critical insight into functional composition changes of this nature.

*Future Work*

Our results illustrates a well known but increasingly pervasive problem in synthesis-based ecology of distinguishing true ecological patterns from methodological artifacts.
We present here the conclusion best supported by available data with a number of caveats and suggestions for next steps forward.

The most pressing need is for intentional and targeted data collection efforts.
We join others in the call for increased monitoring in under sampled areas and continued efforts to centralize existing data sources [@gonzalez2016; @hughes2021; @vellend2017]. Data that fill geographic, taxonomic and trait gaps should be prioritized over further collection of data that replicate existing biases. One relatively low-cost high-reward data investment is collation of additional species-level trait means. Intentional trait selection is critical for linking functional patterns to ecological process [@zhu2017], however synthesis is constrained to the traits in a few taxa-specific databases. Prioritization of additional traits linked to environmental change drivers (response traits) or ecosystem function (effect traits) in particular would provide a powerful link to existing frameworks for understanding functional change [@lavorel2002].

>>>>

#### Data

@@

```{r echo = FALSE, results = "asis"}
tibble(Taxa = c("Mammals", "Birds", "Amphibians"),
       Timeseries_Count = c(48, 2380, 11),
       Species_Count = c(184, 700, 184),
       Trait_Source = c("Elton Traits", "Elton Traits", "Amphibio"),
       Traits = c("body mass, diet, active diel period", "body mass, diet, nocturnality, \nforest foraging strata, pelagic specialist", "habitat, diet, active diel period, \nactivity seasonality, body mass, body length, \nmin maturation size, max maturation size, \nmin offspring size, max offspring size, \nreproductive output, breeding strategy")
) %>%
  mutate_all(linebreak, align = "l") %>% 
  knitr::kable(format = "latex", col.names = c("Taxa", "Number of \nTimeseries", "Number of \nSpecies", "Trait Source", "Traits"), escape = FALSE, 
               caption = "Summary of the data in the final trait database.") %>%
  kable_classic() %>% 
  kable_styling(latex_options= c("scale_down", "hold_position")) #%>%
   #kableExtra::column_spec(2, width = "5cm") #%>%
  # kableExtra::column_spec(4, width = "10cm")
```

@@

#### Diversity Metrics

@@

#### Null Models

@@

#### Analysis

@@

#### Evidence of Ecological Processes

@@

#### Potential Methodological Limitations

Here we approach the question of functional change using the best available data and biodiversity synthesis approaches.
However, a number of gaps in best practices may be obscuring a true underlying trend.
First, the Biotime database is not a representative sample of the world's biodiversity or areas of greatest threat [@gonzalez2016; @vellend2017], and the subset of data in this study exhibits a strong Northern Hemisphere bias. We may simply not have data from those areas experiencing the greatest perturbation [@hughes2021]. We do however emphasize that even disturbed communities maintain functional structure [@edwards2013; @matuoka2020], and the inclusion of more heavily impacted sites would not necessarily change the result.

@@

Third, while we use here the most common metrics for describing functional diversity they do not measure some potentially important aspects of the functional space.
Most notably, the metrics we calculated do not capture shifts in the location of the functional space.
For example, two communities could have very similar metric values but no overlap in their trait spaces.
This is especially relevant in the context of biodiversity change as a species loss could be replaced by a species with very different functional attributes, but the replacement would go undetected if the new species expanded the trait space by the same degree and had similar abundance.
This scenario may be common in communities tracking changing environmental conditions.
Approaches for assessing shifts and overlaps in functional space are still relatively new [@barros2016; @blonder2018; @mammola2019] but could shed critical insight into functional composition changes of this nature.

#### Future Work

Our results illustrate a well known but increasingly pervasive problem in synthesis-based ecology of distinguishing true ecological patterns from methodological artifacts.
We present here the conclusion best supported by available data with a number of caveats and suggestions for next steps forward.
The most pressing need is for intentional and targeted data collection efforts.
We join others in the call for increased monitoring in under sampled areas and continued efforts to centralize existing data sources [@gonzalez2016; @hughes2021; @vellend2017]. Data that fill geographic, taxonomic and trait gaps should be prioritized over further collection of data that replicate existing biases. One relatively low-cost high-reward data investment is collation of additional species-level trait means. Intentional trait selection is critical for linking functional patterns to ecological processes [@zhu2017], however synthesis is constrained to the traits in a few taxa-specific databases. Prioritization of additional traits linked to environmental change drivers (response traits) or ecosystem function (effect traits) in particular would provide a powerful link to existing frameworks for understanding functional change [@lavorel2002].

### References
